---
title: "Master Thesis R-Code"
author: "Marc Suter"
date: "2025-01-04"
output:
  html_document: default
  pdf_document: default
---

# Data Processing

### Loading of Data and Packages

Loading of preprocessed Data from Project Server.

```{r}
load("/Volumes/cds-masterprojects/diffeat/data/ diffeat_data.Rdata")
df <- diffeat_df
```

```{r}
library("ggplot2")
library("reshape2")
library("tidyr")
library("dplyr")
library("ppcor")
library("esquisse")
library("pander")
library("car")
library("Hmisc")
```

## Sample Overview (Pre-Exclusion)

```{r}
df %>% distinct(id) %>% nrow()  # Unique participants
range(df$age, na.rm = TRUE)  # Age range
```

### Remove FRK Questionnaire Items

Remove FRK Questionnaire Items, Rename Demographic Variables

```{r}
# Remove FRK Questionnaire Items
df <- df[, !names(df) %in% grep("^v_(300|3[0-1][0-2]?)$", names(df), value = TRUE)]

### Rename Demographic Variables
colnames(df)[colnames(df) == "v_319"] <- "first_language"
colnames(df)[colnames(df) == "v_317"] <- "school_education" 
colnames(df)[colnames(df) == "v_318"] <- "work_education" 
```

## Questionnaire Data

### Exclusion Based on Eating Disorder Diagnosis

Participants with a current diagnosis of an eating disorder (ED) were removed from the sample.

```{r}
# Identify participants with ED
df %>% filter(current_ed == 1) %>% dplyr::select(id, current_ed, current_ed_type) 
df <- df %>% filter(id != 80) # Exclude participant with Bulimic Eating Behavior
#Removal of respective ED-related variables
df <- df[, !grepl("^(current_ed|current_ed_type|past_ed_type|past_ed)$", names(df))] 
```

### Eating Attitudes Test 26 (EAT-26)

Responses outside the expected range (1–7) were identified and recoded to missing.

```{r}
EAT_var_all <- colnames(df)[colnames(df) %in% paste0("v_", c(191:216))]

# Identify values outside valid range (1-7)
df %>%
  dplyr::select(id, EAT_var_all) %>%
  gather(key = "variable", value = "value", EAT_var_all) %>%
  filter(value > 7 | value < 1) %>%
  distinct(id, variable)

#### Identify 
df$v_201[df$id == 94] 
df$v_215[df$id == 45] 

#recode as NA, as 0 does not exsist in the EAT-26
df$v_201[df$id == c(94)] <- NA  
df$v_215[df$id == c(45)] <- NA
```

#### Reverse Coding

Variable v_215, Item 25 was reverse-coded, since it was reverse-worded in the questionnaire ("Es macht mir Spaß, neue ausgiebige Nahrung zu probieren." [“I enjoy trying out new, rich foods.”]).

```{r}
df <- df %>% mutate(v_215 = 7 - v_215)
```

#### Recoding of EAT-26 Items

Items were then scored according to the EAT-26 manual.

```{r}
df[, EAT_var_all] <- lapply(df[, EAT_var_all], function(x) {
  x <- as.numeric(x)
  x[x == 1] <- 0
  x[x == 2] <- 0
  x[x == 3] <- 0
  x[x == 4] <- 1
  x[x == 5] <- 2
  x[x == 6] <- 3
  return(x)
})
```

#### Total Score and Cutoff-Based Exclusion

Identify participants above EAT-26 cutoff (\>=20), remove specific participants (id = 38, 88, 111) above cutoff.

```{r}
df$EAT_26_total <- rowSums(df[EAT_var_all], na.rm = TRUE)
# Identify participants above EAT-26 cutoff (>=20)
EAT_26_cut_off <- df %>% filter(EAT_26_total >= 20) %>% distinct(id, .keep_all = TRUE)
df <- df %>% filter(!id %in% c(38, 88, 111)) 
rm(EAT_var_all, EAT_26_cut_off)
```

## FEV Subscales ("Fragebogen zum Essverhalten")

#### Subscale "Kognitive Kontrolle" [cognitive control]

```{r}
FEV_var_cognitive_control <- c("v_220", "v_222", "v_226","v_230", "v_234", "v_237", "v_239", "v_244","v_246", "v_248", "v_249", "v_251","v_337", "v_339", "v_340", "v_342", "v_344", "v_347", "v_348", "v_253", "v_254")
```

#### Subscale "Störbarkeit des Essverhaltens" [Disruptability]

```{r}
FEV_var_disruptability <- c("v_217", "v_218","v_223", "v_225", "v_227", "v_229", "v_231", "v_232", "v_236", "v_241", "v_243", "v_247", "v_252", "v_341", "v_345", "v_346")
```

#### Subscale "Hunger" [hunger]

```{r}
FEV_var_hunger <- c("v_219","v_221", "v_224", "v_228", "v_233", "v_235", "v_238","v_240", "v_242", "v_245", "v_250", "v_336", "v_338", "v_343")
```

The rest of the questionnaire was not used for the analysis, as they corresponded to no subscale.

```{r}
FEV_rest <- colnames(df)[colnames(df) %in% paste0("v_", c(349:357))] 
cols_to_keep <- c(names(df)[!names(df) %in% FEV_rest], FEV_var_cognitive_control, FEV_var_hunger)
df <- df[, cols_to_keep]
```

### Data Management of FEV Items {.tabset}

#### Check for invalid values outside of the range of the questionnaire

Responses outside the expected range (1–4) were identified and recoded to missing.

```{r}
EAT_var_all <- colnames(df)[colnames(df) %in% paste0("v_", c(191:216))]

# Identify values outside valid range (1-4)
invalid_values <- df %>% 
  dplyr::select(id, all_of(FEV_var_disruptability)) %>% 
  pivot_longer(cols = all_of(FEV_var_disruptability), names_to = "variable", values_to = "value") %>% 
  filter(value < 1 | value > 4) %>% 
  distinct(id, variable)

# Identify affected participants and items
print(invalid_values)
df$v_252[df$id == 29]
df$v_247[df$id == 81]
df$v_252[df$id == 82]
df$v_241[df$id == 105]
df$v_241[df$id == 113]
df$v_217[df$id == 122]
```

Some items contained invalid values; however, these were coded as 0, which is not a valid response option in the EAT-26. Therefore, they were later recoded as missing (NA).

Before that, corresponding IDs of naturally missings were identified in FEV_var_disruptability.

```{r}
missing_info <- which(is.na(df[FEV_var_disruptability]), arr.ind = TRUE)  
missing_data <- data.frame(  
  id = df$id[missing_info[, 1]],  # Get   
  Column = colnames(df[FEV_var_disruptability])[missing_info[, 2]]  
)  

missing_summary <- missing_data %>%
  dplyr::group_by(id, Column) %>%  
  dplyr::summarise(Missing_Count = n(), .groups = "drop") 
print(missing_summary)
```

There were some missing values in the FEV_disruptability items; however, none of the items were entirely missing. Specifically, five participants had missing values in the items v_341, v_345, and v_346.

#### Recoding

Recoding (values \< 1): All values equal to 0 in the FEV_disruptability items were recoded as missing (NA). This step is essential prior to converting the study coding scheme (1, 2) into the questionnaire coding format (0, 1).

```{r}
df[FEV_var_disruptability] <- lapply(df[FEV_var_disruptability], function(x) {  
  x <- as.numeric(x)  
  x[x == 0] <- NA  
  return(x)  
})

```

#### Reverse Coding of Items

```{r}
df <- df %>% mutate_at(vars(matches("^v_247$|^v_241$|^v_232$")), ~3 - .)
```

#### Dichotomization

The response values for items 39, 33, and 24 (v_341, v_345, v_346) were recoded such that responses 1 and 2 were coded as ‘1’, and responses 3 and 4 were coded as ‘0’.

```{r}
recoding_1to4 <- colnames(df)[colnames(df) %in% c("v_341", "v_345", "v_346")]

for (col_name in recoding_1to4) {
  df[[col_name]] <- as.character(df[[col_name]])
  df[[col_name]][df[[col_name]] %in% c("1", "2")] <- "1"
  df[[col_name]][df[[col_name]] %in% c("3", "4")] <- "0"
  df[[col_name]] <- as.numeric(df[[col_name]])
}

  df[[col_name]] <- as.numeric(df[[col_name]]) # Convert back to numeric
```

#### Final Recoding

Subsequently, the study coding was converted to the questionnaire coding format, where 1 remained 1 and 2 was recoded as 0.

```{r}
df[FEV_var_disruptability] <- lapply(df[FEV_var_disruptability], function(x) {
  x <- as.numeric(x)
  x[x == 2] <- 0
  return(x)
})
```

#### FEV Disruptability Score Calculation

```{r}
df[FEV_var_disruptability] <- lapply(df[FEV_var_disruptability], as.numeric)
df$FEV_disruptability <- rowMeans(df[, FEV_var_disruptability], na.rm = TRUE)
```

#### Handling of Outliers

A custom function was used to detect outliers defined as values beyond three standard deviations from the mean. Based on this, one outlier (ID = 12) in FEV_disruptability was identified and removed from the dataset.

```{r}
detect_outliers <- function(variable, id_column) { 
  mean_score <- mean(variable, na.rm = TRUE)  
  sd_score <- sd(variable, na.rm = TRUE)  
  lower_bound <- mean_score - 3 * sd_score  
  upper_bound <- mean_score + 3 * sd_score  
  outlier_ids <- unique(id_column[!(variable >= lower_bound & variable <= upper_bound)])  
  print(paste("Id(s):", toString(outlier_ids)))  
}
detect_outliers(df$FEV_disruptability, df$id) 
df <- df %>% filter(id != 12)
```

#### Clear Environment

```{r}
df <- df[, !names(df) %in% c(FEV_var_disruptability)] #
rm(list = setdiff(ls(), c("diffeat_df", "df","detect_outliers")))
```

## BIS-11

### Merge BIS-11 Data

```{r}
load("/Volumes/cds-masterprojects/diffeat/data/diffeat_bis_data.Rdata")
bis_data_sub <- bis_data_sub %>% rename(BIS_attention = BARatt, BIS_motor = BARmot, BIS_nonplanning = BARnpl)
df <- merge(df, bis_data_sub, by = "id")
```

### Handle Missing Data

```{r}
missing_ids <- df %>% dplyr::select(id, BIS_attention, BIS_motor, BIS_nonplanning) %>% pivot_longer(cols = -id, names_to = "measure", values_to = "value") %>% filter(is.na(value))
print(missing_ids)
rm(missing_ids)
```

There were participants (id = 15, 22), that will be removed in the affected analysis.

### Handle Outliers

The custom function from above was used again to detect outliers based on set criteria (3 standard deviations from the mean).

```{r}
detect_outliers(df$BIS_attention, df$id)
detect_outliers(df$BIS_motor, df$id)
detect_outliers(df$BIS_nonplanning, df$id) 
```

In *BIS_attention* and *BIS_nonplanning* no outliers were detected. In *BIS_motor*, one outlier (id = 96) was detected.

### Compute BIS Total Score

```{r}
df <- df %>% mutate(BIS_total = rowSums(dplyr::select(., BIS_attention, BIS_motor, BIS_nonplanning), na.rm = FALSE))
rm(bis_data_sub)
rm(list = setdiff(ls(), c("diffeat_df", "df","detect_outliers"))) #clear environment
```

## UPPS

```{r}
UPPS_var_urgency <- c("v_256", "v_260", "v_264", "v_268", "v_272", "v_278", "v_282", "v_286", "v_290", "v_295", "v_297", "v_299")

UPPS_var_sensation_seeking <- c("v_257", "v_261", "v_265", "v_269", "v_273", "v_275", "v_279", "v_283", "v_287", "v_291", "v_296", "v_298")

UPPS_var_perseverance <- c("v_258", "v_262", "v_266", "v_270", "v_274", "v_276", "v_280", "v_284", "v_288", "v_292")

UPPS_var_premeditation <- c("v_255", "v_259", "v_263", "v_267", "v_271", "v_277", "v_281", "v_285", "v_289", "v_293", "v_294")
```

### UPPS reverse coding

#v_262 and v_292 were already recoded and v_297 was falsely recoded during preprocessing and must be recoded again.

```{r}
variables_to_reverse <- c("v_256", "v_257", "v_260", "v_261", "v_264", "v_265", 
                          "v_268", "v_269", "v_272", "v_273", "v_275", "v_278", "v_279", 
                          "v_282", "v_283", "v_286", "v_287", "v_290", "v_291", 
                          "v_295", "v_296", "v_297", "v_298", "v_299")

for (var in variables_to_reverse) {
  df[[var]] <- 5 - df[[var]]}
```

### UPPS subscale score calculation

```{r}
df$UPPS_urgency <- rowMeans(df[, UPPS_var_urgency], na.rm = TRUE)  
df$UPPS_sensation_seeking <- rowMeans(df[, UPPS_var_sensation_seeking], na.rm = TRUE) 
df$UPPS_perseverance <- rowMeans(df[, UPPS_var_perseverance], na.rm = TRUE)
df$UPPS_premeditation <- rowMeans(df[, UPPS_var_premeditation], na.rm = TRUE) 
df$UPPS_total <- rowMeans(df[, c(UPPS_var_urgency, UPPS_var_sensation_seeking, UPPS_var_perseverance, UPPS_var_premeditation)], na.rm = TRUE) 
```

### Check for NAs

```{r}
any(is.na(df$UPPS_urgency))
any(is.na(df$UPPS_sensation_seeking))
any(is.na(df$UPPS_perseverance)) 
any(is.na(df$UPPS_premeditation))
```

### Handle Outliers

One outlier was detected in UPPS_urgency (id = 49) and will be removed before the respective analysis.

```{r}
detect_outliers(df$UPPS_urgency, df$id) 
detect_outliers(df$UPPS_sensation_seeking, df$id)
detect_outliers(df$UPPS_perseverance, df$id)
detect_outliers(df$UPPS_premeditation, df$id)
```

```{r}
df <- df[, !names(df) %in% c(UPPS_var_urgency, UPPS_var_sensation_seeking, UPPS_var_perseverance, UPPS_var_premeditation)]
```

## Tractography Data

### Data Filtering and Processing

#### Select Relevant Measures

Filter out the superior vetegemental area (sVTA) and nucleus accumbens (NAcc) tract, as there was no impulsvity-association found (Tisdall, 2024) and select only fractional anisotropy (FA) and radial diffusivity (iRD) measures.

```{r}
df <- df %>% filter(measure %in% c("FA", "iRD"))
df <- df %>% filter(!tract %in% c("sVTA-NAcc (L)", "sVTA-NAcc (R)", "sVTA-NAcc (bil)"))
```

#### Extract Hemisphere Information

```{r}
df <- df %>% mutate(
  hemisphere = case_when(grepl("\\(L\\)", tract) ~ "L", grepl("\\(R\\)", tract) ~ "R", TRUE ~ "B"),
  tract = gsub("\\(L\\)|\\(R\\)|\\(bil\\)", "", tract),
  tract = trimws(tract)
)

df$tract <- as.factor(df$tract)
df$tract <- relevel(df$tract, "MPFC-NAcc")
df$hemisphere <- factor(df$hemisphere,
                        levels = c("L", "B", "R"),
                        labels = c("Left", "Bilateral", "Right"))

df$hemisphere <- as.factor(df$hemisphere)
df$hemisphere <- relevel(df$hemisphere, "Left")

levels(df$hemisphere)
```

#### Rename Numeric Columns

```{r}
numeric_cols <- grep("^\\d+$", names(df))
names(df)[numeric_cols] <- paste0("node_", seq_along(numeric_cols))
rm(numeric_cols)
```

### Preprocessing

create middle 50% incidies

```{r}
nodes <- paste0("node_", 26:75)
any(is.na(df[, nodes])) #NA values check in the middle 50% of the tracts     
df$mean_middle_50 <- rowMeans(df[, nodes]) 
```

### FA Plot

```{r}
df_FA <- df %>%
  filter(measure == "FA") %>%
  gather(key = "nodes", value = "value", node_26:node_75)   

df_FA$nodes <- gsub("node_", "", df_FA$nodes)
df_FA$id <- as.factor(df_FA$id)
df_FA$nodes <- as.numeric(df_FA$nodes)
df_FA$value <- as.numeric(df_FA$value)

FA_Plot <- ggplot(data = df_FA, 
                  mapping = aes(x = nodes, y = value, color = as.factor(id), group = id)) +
  geom_line(size = 0.25, alpha = 0.5) +  
  facet_grid(hemisphere ~ tract) +
  ggtitle("FA Values") +
  theme_bw() +
  guides(color = "none") +
  scale_x_continuous(breaks = c(30, 40, 50, 60, 70)) +
  xlab("Nodes") +  
  ylab("Value") +  
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_blank(),  
    strip.text.x = element_text(size = 12),  
    strip.text.y = element_text(size = 12),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10, margin = margin(t = 2, r = 0, b = 0, l = 0)),  
    axis.text.y = element_text(size = 10, margin = margin(t = 0, r = 2, b = 0, l = 0))  
  )

print(FA_Plot)

ggsave("FA_middle50.png", plot = FA_Plot, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300) 
```

### iRD Plot

```{r}
df_iRD <- df %>%
  filter(measure == "iRD") %>%
  gather(key = "nodes", value = "value", node_26:node_75)   

df_iRD$nodes <- gsub("node_", "", df_iRD$nodes)
df_iRD$id <- as.factor(df_iRD$id)
df_iRD$nodes <- as.numeric(df_iRD$nodes)
df_iRD$value <- as.numeric(df_iRD$value)

iRD_Plot <- ggplot(data = df_iRD, 
                   mapping = aes(x = nodes, y = value, color = as.factor(id), group = id)) +
  geom_line(size = 0.25, alpha = 0.5) +  
  facet_grid(hemisphere ~ tract) +
  ggtitle("iRD Values") +
  theme_bw() +
  guides(color = "none") +
  scale_x_continuous(breaks = c(30, 40, 50, 60, 70)) +
  xlab("Nodes") +  
  ylab("Value") +  
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_blank(),  
    strip.text.x = element_text(size = 12),  
    strip.text.y = element_text(size = 12),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10, margin = margin(t = 2, r = 0, b = 0, l = 0)),  
    axis.text.y = element_text(size = 10, margin = margin(t = 0, r = 2, b = 0, l = 0))  
  )

print(iRD_Plot)

ggsave("iRD_middle50.png", plot = iRD_Plot, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)

```

### 

Save the data on the respective server.

# Data Analysis

## Descriptive Statistics

### Demographic Overview

```{r}
#Final sample size
nrow(df)/24  

#Mean age
round(mean(df$age),1)  

#Age standard deviation
round(sd(df$age),1)   

# Gender proportions
gender <- table(df$gender) / 24
round((table(df$gender)/ 24)/sum(table(df$gender)/ 24),3)*100  

#School education proportions
table(df$school_education) / 24
round((table(df$school_education)/ 24)/sum(table(df$school_education)/ 24),3)*100



```

### Language Data Cleaning

```{r}
df$first_language <- as.character(df$first_language)
df$first_language[df$first_language == "deutsch"] <- "German"
df$first_language[df$first_language == "Deutsch"] <- "German"
df$first_language[df$first_language == "DEUTSCH"] <- "German"
df$first_language[df$first_language == "deutsch englisch"] <- "German/English"
df$first_language[df$first_language == "Englisch"] <- "English"
df$first_language[df$first_language == "Russisch"] <- "Russian"
df$first_language[df$first_language == "Spanisch"] <- "Spanish"
df$first_language[df$first_language == "Deutsch und Russisch"] <- "German/Russian"

table(df$first_language) / 24
round((table(df$first_language)/ 24)/sum(table(df$first_language)/ 24),3)*100 #Percentage
```

### Body Mass Index (BMI)

```{r}
df$BMI <- df$weight_kg / (df$height_m^2)
mean(df$BMI)
sd(df$BMI)
df <- df %>% dplyr::select(-weight_kg, -height_m)
```

# Hypothesis Testing

## Hypothesis 1a: FA of iVTA-NAcc Bilateral with UE

```{r}
H1a <- subset(df, tract == "iVTA-NAcc" & measure == "FA" & hemisphere == "Bilateral")
df <- df %>% filter(!id %in% c(12, 38, 88, 111))
nrow(df)/24
```

### Variable Standardization and Normality Checks

```{r}
H1a$mean_middle_50_sdt <- scale(H1a$mean_middle_50)
H1a$FEV_disruptability_sdt <- scale(H1a$FEV_disruptability)

shapiro.test(H1a$FEV_disruptability_sdt)
shapiro.test(H1a$mean_middle_50_sdt)
```

### Partial Correlation Analysis

```{r}
H1a_result <- pcor.test(H1a$mean_middle_50_sdt, 
                        H1a$FEV_disruptability_sdt, 
                        cbind(H1a$gender, H1a$age, H1a$BMI))
H1a_result
explained_variance <- (H1a_result$estimate)^2
```

### Scatterplot for Hypothesis 1a

```{r}
residuals_mean <- residuals(lm(mean_middle_50_sdt ~ gender + age + BMI, data = H1a))
residuals_fev <- residuals(lm(FEV_disruptability_sdt ~ gender + age + BMI, data = H1a))

data <- data.frame(residuals_mean, residuals_fev)

H1a_scatter <- ggplot(data, aes(x = residuals_mean, y = residuals_fev)) +
  geom_point(color = "#508F81", size = 3) +
  geom_smooth(method = "lm", color = "#334F49", linewidth = 1.2, se = FALSE) +
  labs(
    x = "Residuals of Middle 50% of VTA-NAcc tract",
    y = "Residuals of Uncontrolled Eating"
  ) +
  theme_bw() +
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 14),
    panel.border = element_rect(color = "black", fill = NA),
    plot.margin = margin(1, 1, 1, 1, "cm")
  )

ggsave("H1a_scatter.png", plot = H1a_scatter, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)
```

### BMI Distribution in Hypothesis 1a

```{r}
hist(H1a$BMI, main = "", xlab = "BMI", ylab = "Frequency", col = "#508F81", border = "black", breaks = 30, font.lab = 2, cex.lab = 1.5, cex.axis = 1.2, bty = "l")
abline(v = 25, col = "red", lwd = 2, lty = 2)
abline(v = 18.5, col = "blue", lwd = 2, lty = 2)
legend("topright", legend = c("Overweight (≥25)", "Underweight (≤18.5)"), col = c("red", "blue"), lty = 2, lwd = 2, bty = "o", cex = 0.8, inset = 0.05)

sum(H1a$BMI >= 25)
sum(H1a$BMI <= 18.5)
```

```{r}
rm(list = setdiff(ls(), c("diffeat_df", "df")))
```

## Exploratory Analyses

### Exploratory Node-Wise Analysis

```{r}
Exp_1 <- df #The full dataframe is used.
nrow(Exp_1)/24 # Exp_1 sample size
unique(Exp_1$hemisphere)
unique(Exp_1$tract)
unique(Exp_1$measure)
```

#### Colors for each tract

```{r}
colors <- list(mpfc = rgb(15, 209, 219, maxColorValue = 255), ains = rgb(226, 24, 2, maxColorValue = 255),
               amyg = rgb(0, 0, 139, maxColorValue = 255),  ivta = rgb(255, 215, 0, maxColorValue = 255))
```

#### Prepare the data for FA regression

```{r}
Exp_1_fa_long <- df %>%
  filter(measure == "FA") %>%
  pivot_longer(cols = starts_with("node_"), 
               names_to = "nodes", 
               values_to = "value") %>%
  mutate(nodes = as.numeric(gsub("node_", "", nodes)))
```

```{r}
tracts <- c("MPFC-NAcc", "AIns-NAcc", "Amy-NAcc", "iVTA-NAcc")
hemispheres <- c("Left", "Bilateral", "Right")
```

```{r}
# Compute regression results
Exp_1_fa_regs <- Exp_1_fa_long %>%
  filter(tract %in% tracts) %>%
  group_by(tract, hemisphere, nodes) %>%
  filter(!(tract == "MPFC-NAcc" & id == '10')) %>%   # Remove outlier for MPFC-NAcc, reported from Supervisor can
  do({
    reg <- lm(FEV_disruptability ~ value, data = .)
    tval <- summary(reg)$coefficients[2,3]
    pval <- summary(reg)$coefficients[2,4]
    data.frame(TVALUE = tval, PVALUE = pval)
  }) %>%
  ungroup()
```

```{r}
Exp_1_fa_regs <- Exp_1_fa_regs %>%
  mutate(HEMISPHERE = factor(hemisphere, levels = hemispheres),
         TRACT = factor(tract, levels = tracts))
```

#### Plotting of results for the t Values of FA

```{r}
tvals_FA <- ggplot(Exp_1_fa_regs, aes(x = nodes, y = TVALUE, color = TRACT, group = TRACT)) + 
  geom_point(size = 0.5, shape=21) +
  facet_grid(HEMISPHERE ~ TRACT) +
  ggtitle(expression(paste("FA ", italic("t"), "-Values"))) +  
  theme_bw() +
  guides(color = "none") +
  xlab("Nodes") +  
  ylab(expression(paste(italic("t"), " Values"))) +  
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_blank(),  
    strip.text.x = element_text(size = 12),  
    strip.text.y = element_text(size = 12),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10, margin = margin(t = 2)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 2))  
  ) +
  scale_color_manual(values = c("MPFC-NAcc" = colors$mpfc, 
                                "AIns-NAcc" = colors$ains, 
                                "Amy-NAcc" = colors$amyg, 
                                "iVTA-NAcc" = colors$ivta)) +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed", size = 0.5)
```

```{r}
print(tvals_FA)
```

```{r}
ggsave("tvals_FA", plot = tvals_FA, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)
```


#### Plotting of results for the p Values of FA


```{r}
pvals_FA <- ggplot(Exp_1_fa_regs, aes(x = nodes, y = PVALUE, color = TRACT, group = TRACT)) + 
  geom_point(size = 0.5, shape=21) +
  facet_grid(HEMISPHERE ~ TRACT) +
  ggtitle(expression(paste("FA ", italic("p"), "-Values"))) + 
  theme_bw() +
  guides(color = "none") +
  xlab("Nodes") +  
  ylab(expression(paste(italic("p"), " Values"))) + 
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_blank(),  
    strip.text.x = element_text(size = 12),  
    strip.text.y = element_text(size = 12),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10, margin = margin(t = 2)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 2))  
  ) +
  scale_color_manual(values = c("MPFC-NAcc" = colors$mpfc, 
                                "AIns-NAcc" = colors$ains, 
                                "Amy-NAcc" = colors$amyg, 
                                "iVTA-NAcc" = colors$ivta)) +
  geom_hline(yintercept = 0.05, color = "grey", linetype = "dashed", size = 0.5)
```

```{r}
print(pvals_FA)
```

```{r}
ggsave("pvals_FA", plot = pvals_FA, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)
```

#### Prepare the data for iRD regression

```{r}
Exp_1_iRD_long <- df %>%
  filter(measure == "iRD") %>%
  pivot_longer(cols = starts_with("node_"), 
               names_to = "nodes", 
               values_to = "value") %>%
  mutate(nodes = as.numeric(gsub("node_", "", nodes)))
```

```{r}
tracts <- c("MPFC-NAcc", "AIns-NAcc", "Amy-NAcc", "iVTA-NAcc")
hemispheres <- c("Left", "Bilateral", "Right")
```

```{r}
Exp_1_iRD_regs <- Exp_1_iRD_long %>%
  filter(tract %in% tracts) %>%
  group_by(tract, hemisphere, nodes) %>%
  filter(!(tract == "MPFC-NAcc" & id == '10')) %>%   # Remove outlier for MPFC-NAcc
  do({
    reg <- lm(FEV_disruptability ~ value, data = .)
    tval <- summary(reg)$coefficients[2,3]
    pval <- summary(reg)$coefficients[2,4]
    data.frame(TVALUE = tval, PVALUE = pval)
  }) %>%
  ungroup()
```

```{r}
Exp_1_iRD_regs <- Exp_1_iRD_regs %>%
  mutate(HEMISPHERE = factor(hemisphere, levels = hemispheres),
         TRACT = factor(tract, levels = tracts))
```


#### Plotting of results for the t Values of iRD


```{r}
tvals_iRD <- ggplot(Exp_1_iRD_regs, aes(x = nodes, y = TVALUE, color = TRACT, group = TRACT)) + 
  geom_point(size = 0.5, shape=21) +
  facet_grid(HEMISPHERE ~ TRACT) +
  ggtitle(expression(paste("iRD ", italic("t"), "-Values"))) +
  theme_bw() +
  guides(color = "none") +
  xlab("Nodes") +  
  ylab(expression(paste(italic("t"), " Values"))) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_blank(),  
    strip.text.x = element_text(size = 12),  
    strip.text.y = element_text(size = 12),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10, margin = margin(t = 2)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 2))  
  ) +
  scale_color_manual(values = c("MPFC-NAcc" = colors$mpfc, 
                                "AIns-NAcc" = colors$ains, 
                                "Amy-NAcc" = colors$amyg, 
                                "iVTA-NAcc" = colors$ivta)) +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed", size = 0.5)
```

```{r}
print(tvals_iRD)
```

```{r}
ggsave("tvals_iRD.png", plot = tvals_iRD, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)
```

#### Plotting of results for the p Values of iRD

```{r}
pvals_iRD <- ggplot(Exp_1_iRD_regs, aes(x = nodes, y = PVALUE, color = TRACT, group = TRACT)) + 
  geom_point(size = 0.5, shape=21) +
  facet_grid(HEMISPHERE ~ TRACT) +
  ggtitle(expression(paste("iRD ", italic("p"), "-Values"))) + 
  theme_bw() +
  guides(color = "none") +
  xlab("Nodes") +  
  ylab(expression(paste(italic("p"), " Values"))) + 
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_blank(),  
    strip.text.x = element_text(size = 12),  
    strip.text.y = element_text(size = 12),  
    axis.title.x = element_text(size = 12),  
    axis.title.y = element_text(size = 12),  
    axis.text.x = element_text(size = 10, margin = margin(t = 2)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 2))  
  ) +
  scale_color_manual(values = c("MPFC-NAcc" = colors$mpfc, 
                                "AIns-NAcc" = colors$ains, 
                                "Amy-NAcc" = colors$amyg, 
                                "iVTA-NAcc" = colors$ivta)) +
  geom_hline(yintercept = 0.05, color = "grey", linetype = "dashed", size = 0.5)
```

```{r}
print(pvals_iRD)
```

```{r}
ggsave("iRD_pvals.png", plot = pvals_iRD, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)
```


### Exploratory Correlation Analysis of Impulsivity Measures

#### Correlation Matrix Plot

A matrix plot was created with the outliers form the processing step removed (IDs = 15, 22, 71, 70).

```{r}
Exp_2 <- subset(df, tract == "iVTA-NAcc" & measure == "FA" & hemisphere == "Bilateral")

Exp_2 <- Exp_2[!Exp_2$id %in% c(15, 22, 71, 70),]

scales <- Exp_2[, c("BIS_attention", "BIS_motor", "BIS_nonplanning", "BIS_total", 
                   "UPPS_urgency", "UPPS_sensation_seeking", "UPPS_perseverance", 
                   "UPPS_premeditation", "UPPS_total")]

#Compute correlation and p-values
rcorr_res <- rcorr(as.matrix(scales), type = "pearson")
cor_matrix <- rcorr_res$r
p_matrix <- rcorr_res$P

#Convert to long format
cor_data <- melt(cor_matrix)
p_data <- melt(p_matrix)
colnames(cor_data) <- c("Var1", "Var2", "value")
cor_data$p_value <- p_data$value

cor_data <- cor_data %>% mutate(Var1 = factor(Var1, levels = colnames(scales)), Var2 = factor(Var2, levels = colnames(scales))) %>% filter(as.numeric(Var1) > as.numeric(Var2))

#Add stars based on p-value
  cor_data$stars <- cut(cor_data$p_value,
                        breaks = c(-Inf, 0.01, 0.05, Inf),
                        labels = c("**", "*", ""))
  cor_data$label <- paste0(round(cor_data$value, 2), cor_data$stars)
  
full_grid <- expand.grid(Var1 = colnames(scales), Var2 = colnames(scales)) %>%
  mutate(Var1 = factor(Var1, levels = colnames(scales)), Var2 = factor(Var2, levels = colnames(scales)))

plot_data <- left_join(full_grid, cor_data, by = c("Var1", "Var2"))

#New labels for better readability
new_labels <- c(
  BIS_attention = "Attention (BIS-11)",
  BIS_motor = "Motor (BIS-11)",
  BIS_nonplanning = "Non-Planning (BIS-11)",
  BIS_total = "BIS Total",
  UPPS_urgency = "Urgency (UPPS)",
  UPPS_sensation_seeking = "Sensation Seeking (UPPS)",
  UPPS_perseverance = "Perseverance (UPPS)",
  UPPS_premeditation = "Premeditation (UPPS)",
  UPPS_total = "UPPS Total"
)


Exp2_CorPlot <- ggplot(plot_data, aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = label), color = "black", size = 3, na.rm = TRUE) +
  scale_fill_gradient2(low = "#FAE20C", high = "#6E49FA", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", na.value = "white") +
  scale_x_discrete(labels = new_labels) +
  scale_y_discrete(labels = new_labels) +
  coord_fixed() +
  theme_bw(base_size = 12) +
  labs(x = NULL, y = NULL, title = expression(bold("Visual Correlation Matrix"))) +
  theme(plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        panel.grid = element_blank(),
        axis.text.x = element_text(size = 11, angle = 35, vjust = 1, hjust = 1, family = "Arial", color = "black"),
        axis.text.y = element_text(size = 11, family = "Arial", color = "black"),
        legend.position = "none")

print(Exp2_CorPlot)
```

ggsave("Exp2_CorPlot.png", plot = Exp2_CorPlot, device = "png", width = 500/85, height = 450/85, units = "in", dpi = 300)
